---
title: "IST687 Final Project"
output:
  word_document: default
---

```{r setup, include=FALSE}
install.packages("pscl")
install.packages("rsconnect")
install.packages("shiny")
install.packages("randomForest")
library(tidyverse);library(arules);library(arulesViz);library(imputeTS);library(ggplot2);library(plotly);
library(ggmap);library(mapproj);library(rworldmap);library(kernlab);library(caret);library(pscl);library(lmtest);
library(rsconnect);library(shiny); library(randomForest)
```

## Importing Data

```{r}
dataFile <- "https://intro-datascience.s3.us-east-2.amazonaws.com/Resort01.csv"

df <- read_csv(dataFile, show_col_types = FALSE)
head(df)
```

##Adding calculated fields, normalizing data

```{r}
#creating group size category from Adults, Children, and Babies
df$Adults[df$Adults==0 & df$Children==0 & df$Babies==0] <- NA
df$Children[df$Adults==0 & df$Children==0 & df$Babies==0] <- NA
df$Babies[df$Adults==0 & df$Children==0 & df$Babies==0] <- NA

df$Adults <- na_interpolation(df$Adults)
df$Children <- na_interpolation(df$Children)
df$Babies <- na_interpolation(df$Babies)

df$GroupSize <- df$Adults + df$Children + df$Babies
df$GroupSizeCat <- ifelse(df$GroupSize >4, ">4", "<=4")  

#creating stay length from StaysInWeekNights and StaysInWeekendNights
df$StayLength <- df$StaysInWeekNights + df$StaysInWeekendNights
df$StayLength[df$StayLength==0] <- NA

df$StayLength <- ceiling(na_interpolation(df$StayLength))
df$StayLengthCat <- ifelse(df$StayLength  >3, ">3", "<=3")

df$GotWantedRoom <- ifelse(df$ReservedRoomType == df$AssignedRoomType, 1,0)
df$NonRefundable <- ifelse(df$DepositType == "Non Refund", 1,0)

#per the documentation Undefined and SC are the same
df$Meal <- ifelse(df$Meal=="Undefined","SC",df$Meal)

dfNumeric <- df %>% select(where(is.numeric))
dfAssoc <- df %>% mutate_if(is.numeric,as.character)
```

## Summary statistics for all columns

```{r}
summary.data.frame(dfNumeric)
```

## Association rules

```{r}

dfTransactions <- as(dfAssoc, "transactions")

head(itemFrequency(dfTransactions))
itemFrequencyPlot(dfTransactions, topN = 25)

inspect(dfTransactions[1:10])

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.15, conf=0.5), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))


plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:10])
```
## Various Association rules attempts
All non-numeric columns

```{r}

dfTransactions <- as(dfAssoc[c(1,8,9,10,14,15,17,18)], "transactions")

itemFrequencyPlot(dfTransactions, topN = 25)

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.03, conf=0.8), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))
plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:10])
```

Group size, stay length (numerical)
```{r}

dfTransactions <- as(dfAssoc[c(1,21,22)], "transactions")

itemFrequencyPlot(dfTransactions, topN = 25)

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.00001, conf=0.5), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))
plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:10])
```

Group size, stay length (categorical)
```{r}

dfTransactions <- as(dfAssoc[c(1,23,24)], "transactions")

itemFrequencyPlot(dfTransactions, topN = 25)

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.000001, conf=0.01), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))
plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:9])
```

##Plots

```{r}


#Scatter Plots


# Is there a relation between customer type , repeated guest and cancellation?
g1 <- ggplot(df)+
aes(x=IsCanceled, y=IsRepeatedGuest, color=CustomerType) +
geom_point()
g1

# Is there a relation between Market Segment , repeated guest and cancellation?
g2 <- ggplot(df)+
aes(x=IsCanceled, y=IsRepeatedGuest , color= MarketSegment ) +
geom_point()
g2

# Is there a relation between customer type , repeated guest and previous cancellation?
g3 <- ggplot(df)+
aes(x=PreviousCancellations, y=IsRepeatedGuest , color= CustomerType ) +
geom_point()
g3

# Plots

#Room preference by customer type
ggplot(data=df,aes(CustomerType,fill=ReservedRoomType)) +geom_bar(stat="count",position = position_dodge())

#Room preference by market segment
ggplot(data=df,aes(MarketSegment,fill=ReservedRoomType)) +geom_bar(stat="count",position = position_dodge())

# Reservation based on Meal
ggplot(data = df,aes( x = Meal,fill = Meal,y = prop.table(stat(count)),
                          label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
  geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) + coord_flip() +labs(title = "Bookings by Meal", x = "Meal type",y = "Number of bookings") +
  geom_text(stat = "count", aes(label = ..count..), hjust = 5) 


#Booking based on market segment
ggplot(data = df,aes( x = MarketSegment,fill = MarketSegment,y = prop.table(stat(count)),
                          label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
  geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) + coord_flip() +labs(title = "Bookings by Market Segment", x = "MS type",y = "Number of bookings") +
  geom_text(stat = "count", aes(label = ..count..), hjust = 1) 


#Customer Type and Booking
ggplot(data = df,aes( x = CustomerType,fill = CustomerType,y = prop.table(stat(count)),
                          label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
  geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) + coord_flip() +labs(title = "Booking by Customer Type", x = "Customer type",y = "Number of bookings") +
  geom_text(stat = "count", aes(label = ..count..), hjust = 1) 


#Cancellation and Booking
ggplot(data = df,aes( x =IsCanceled,fill = IsCanceled,y = prop.table(stat(count)),
                          label = scales::percent(prop.table(stat(count))))) +  geom_bar() + 
  geom_text(stat = "count", position = position_dodge(1),vjust = 1, hjust=0,size = 3)+scale_y_continuous(labels = scales::percent) + coord_flip() +labs(title = "Booking by Cancellation", x = "Cancellation",y = "Number of bookings") +
  geom_text(stat = "count", aes(label = ..count..), hjust = 1) 

## # Interactive Bar Plot for country
country_table <- df %>%
  count(Country)

country_table %>%
  mutate(Country = fct_reorder(Country, n, .desc = TRUE)) %>%
  plot_ly(x = ~Country, y = ~n) %>% 
  add_bars()  
country_table

#EDA

#Online market segment's cancellation is more
#Online TA	17729	
df%>% group_by(MarketSegment)  %>% summarise(length(IsCanceled))

#Couples booking cancellation is more
#2	31432	
df%>% group_by(df$Adults)  %>% summarise(length(IsCanceled))

#1 night booking cancellations are more
#1	7790	
df%>% group_by(df$StayLength)  %>% summarise(length(IsCanceled))

#'A' type room cancellation is higher
#A	23399	
df %>% group_by(ReservedRoomType)  %>% summarise(length(IsCanceled))

#'A' type room cancellation is higher
#A	17046	
df %>% group_by(df$AssignedRoomType)  %>% summarise(length(IsCanceled))

#customer with no deposit cancellation is higher
#No Deposit	38199	
df %>% group_by(df$DepositType)  %>% summarise(length(IsCanceled))

#Non repeated guests has more cancellations
#0	38282	
df %>% group_by(df$IsRepeatedGuest)  %>% summarise(length(IsCanceled))

visitedMap <- joinCountryData2Map(df,
joinCode = "ISO3",
nameJoinColumn = "Country",
nameCountryColumn = "Country",
mapResolution="coarse" ,
verbose = FALSE
)
par(mai=c(0,0,0.2,0),xaxs="i",yaxs="i")
mapParams <- mapCountryData(visitedMap,
nameColumnToPlot="Country",
oceanCol = "azure2",
missingCountryCol = gray(.8),
addLegend = F,
mapTitle = "Hotel Cancelations Across Countries",
catMethod='categorical')


```

#Previous cancellations, repeated guests, deposit type, group size, stay length
```{r}

dfTransactions <- as(dfAssoc[c(1,12,17,23,24)], "transactions")

itemFrequencyPlot(dfTransactions, topN = 25)

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.006, conf=0.5), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))
plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:10])
```

#Market segment, deposit type, customer type
```{r}
dfTransactions <- as(dfAssoc[c(1,10,17,18)], "transactions")

itemFrequencyPlot(dfTransactions, topN = 25)

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.006, conf=0.5), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))
plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:5])
```

#Babies, children, deposit, group size categorical, stay length categorical
```{r}

dfTransactions <- as(dfAssoc[c(1,6,7,17,23,24)], "transactions")

itemFrequencyPlot(dfTransactions, topN = 25)

ruleset <- apriori(dfTransactions,
                   parameter=list(supp=0.01, conf=0.90), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs",rhs=("IsCanceled=1")))
plot(ruleset)
rulesbysupport <- sort(ruleset, decreasing = TRUE, by = "support")
inspect(rulesbysupport[1:10])
```


##Boxplots

```{r}

for(j in names(dfNumeric)){
  boxplot(df[[j]] ~ df[[1]], main = paste("Boxplot of",j), xlab = "IsCanceled",ylab=j)
}
```

##Barplots

```{r}
for (k in names(df)){
  counts <- table(df[[1]],df[[k]])
  barplot(counts, main = paste("Barplot of",k), xlab = k, 
          col = c("Blue","Orange"), legend = rownames(counts),beside = TRUE)
}
```



#Linear Model
```{r}
mylm<-lm(IsCanceled~LeadTime+StayLength+Adults+Children+IsRepeatedGuest+PreviousCancellations+BookingChanges+RequiredCarParkingSpaces+TotalOfSpecialRequests+GotWantedRoom+NonRefundable,df) 
summary(mylm) #low R squared. Linear model, not good to predict a categorical value
```

#Logistic Regression 1
```{r}
mylogit<-glm(IsCanceled~LeadTime+StayLength+Adults+Children+IsRepeatedGuest+PreviousCancellations+PreviousBookingsNotCanceled+BookingChanges+RequiredCarParkingSpaces+TotalOfSpecialRequests,data=df,family='binomial')
summary(mylogit) 
```


#Logistic Regression 2
```{r}
mylogit2<-glm(IsCanceled~LeadTime+StayLength+Adults+Children+IsRepeatedGuest+PreviousCancellations+PreviousBookingsNotCanceled+BookingChanges+TotalOfSpecialRequests,data=df,family='binomial')
summary(mylogit2)
```


#To determine the better model
```{r}

pR2(mylogit) #look at the McFadden R^2. The 1st logit has a higher psuedo R^2
pR2(mylogit2)

```

#Likelihood Ratio Test
```{r}
lrtest(mylogit,mylogit2) #can reject the null, that the second model has a better fit 
```


##SVM

```{r}
df %>% mutate(IsCanceled = as.factor(IsCanceled)) -> df
trainListHotel <- createDataPartition(y=df$IsCanceled, p=.60, list=FALSE)
trainSetHotel <- data.frame(df[trainListHotel,])
testSetHotel <- data.frame(df[-trainListHotel,])


svm_model <- ksvm(IsCanceled ~ LeadTime + StaysInWeekendNights + StaysInWeekNights + Adults + Children + Babies + IsRepeatedGuest + PreviousCancellations + PreviousBookingsNotCanceled + BookingChanges + RequiredCarParkingSpaces + TotalOfSpecialRequests, data=trainSetHotel,prob.model = TRUE,kernel= "rbfdot", kpar = "automatic",C = 5,cross = 3)

svm_model

svmPred <- predict(svm_model, newdata = testSetHotel,type = "response")
table(testSetHotel$IsCanceled,svmPred)
confusionMatrix(svmPred, testSetHotel$IsCanceled) 

svm_model2 <- ksvm(IsCanceled ~ LeadTime + StaysInWeekendNights + StaysInWeekNights + Adults + Children + Babies + IsRepeatedGuest + PreviousCancellations + PreviousBookingsNotCanceled + BookingChanges + RequiredCarParkingSpaces + TotalOfSpecialRequests + StayLength + GotWantedRoom + NonRefundable + GroupSize, data=trainSetHotel, kpar="automatic",kernel= "rbfdot",prob.model = TRUE,C = 5, cross = 3)

svm_model2

svmPred2 <- predict(svm_model2, newdata = testSetHotel,type = "response")
table(testSetHotel$IsCanceled,svmPred2)
confusionMatrix(svmPred2, testSetHotel$IsCanceled)  
```

#ShinyApp
```{r eval=FALSE, echo=FALSE}

rsconnect::setAccountInfo(name='emanijones', token='3E07F941CA980805DDA56CCD911234BD', secret='o+fiJm1tCIlGVic6GmrqDqL8XRzO9Vypx2j3mVm9')

ui <- fluidPage(
  titlePanel("Probability a Customer Cancels"),
  sidebarLayout(
    sidebarPanel(
      numericInput("LeadTime","Lead Time",0 ),
      numericInput("StayLength","Stay Length",0),
       numericInput("Adults","How Many Adults",0),
      numericInput("Children","How Many Children",0),
      numericInput("IsRepeatedGuest","Is this a Repeated Guest:Yes=1, No=0",0),
      numericInput("PreviousCancellations","How Many Times Has the Customer Cancelled Before",0),
      numericInput("PreviousBookingsNotCanceled","Previous Bookings Not Canceled",0),
      numericInput("BookingChanges", "Number of Booking Changes",0),
      numericInput("RequiredCarParkingSpaces","How Many Parking Spaces Are Needed",0),
      numericInput("TotalOfSpecialRequests","Total Of Special Requests",0)
     
    ),
    mainPanel(textOutput("cancelprob")
      
    )
  )
)

server <- function(input, output, session) {
  preddftry<-predDF<-reactive(data.frame("LeadTime"=input$LeadTime,"StayLength" = input$StayLength,"Adults"=input$Adults,"Children"=input$Children,"IsRepeatedGuest"=input$IsRepeatedGuest,"PreviousCancellations"=input$PreviousCancellations,"PreviousBookingsNotCanceled"=input$PreviousBookingsNotCanceled,"BookingChanges"=input$BookingChanges,"RequiredCarParkingSpaces"=input$RequiredCarParkingSpaces,"TotalOfSpecialRequests"=input$TotalOfSpecialRequests))
  outputprediction <- reactive({predict(mylogit, preddftry(),type='response')})
  output$cancelprob<- renderText(outputprediction())
  
}

shinyApp(ui, server)
```